Why aren't significance tests commonly used for linear regression diagnostics?

### 2022-11-29 paper_comparison/paper_comparison/paper_comparison.Rmd
finish power cmp

<!-- What truly impress us is the huge difference between the estimated power of visual test and the estimated power of F-test. According to the example lineups included in Figure \ref{fig:power-poly-uniform}, at around $log_e(E) = 1$, the  designed visual feature is rarely visible, making the actual data plot indistinguishable from residual plots simulated from the assumed model at this level of difficulty. In fact, zero out of 11 subjects detect the actual data plot positioned at panel 10. However, F-test has more than 60% chance of rejecting the null hypothesis in this case. From a communication perspective, given the fact that the visual difference is unperceivable, the argument that non-linearity present in the fitted model is less convincing to the public. -->

<!-- The margin is largest at around $log_e(E) = 2$. As shown in Figure \ref{fig:power-poly-uniform}, the visual pattern is still undetectable which is similar to $log_e(E) = 1$. The visual test power is close to zero. In contrast, F-test will reject the null hypothesis 90% of the time. -->

<!-- At around $log_e(E) = 3$, the margin gets smaller as the chance of identifying the actual data plot becomes larger. At this level of difficulty, the designed visual features are usually detectable but it may not stand out from the lineup as other null plots may happen to include outliers or visual patterns that are are considered to be more attractive by human, and thus recognized as the most different plot. Without knowing the designed visual features beforehand, it is actually hard to identify the actual data plot by pure image comparison. The corresponding example lineup for $log_e(E) = 3$ shown in Figure \ref{fig:fig:power-poly-uniform} has the actual data plot positioned at panel 20, where two of 11 subjects detect it. It can be observed that a "S-shape" is presented in plot 20, but the signal is not strong enough to attract all 11 subjects, resulting in a visual p-value sightly above $0.1$. At $log_e(E) = 4$, the designed visual features become much clear and attractive in panel 3, leading to a high percentage of rejection of the null hypothesis. Figure \ref{fig:fig:power-poly-uniform} gives example lineups of such cases. It is also interesting to observe that the power curve of visual test behaves similarly to those inappropriate conventional tests until the designed visual features really becomes visible. -->

@@

```{r}
p <- vi_survey %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "Visual test"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
                filter(x_dist == "uniform") %>%
                pivot_longer(F_p_value:SW_p_value) %>%
                mutate(name = gsub("_p_value", " test", name)), 
              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
              method = "glm", 
              method.args = list(family = binomial),
              se = FALSE) +
  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3)  +
  geom_segment(data = data.frame(xi = es), aes(x = xi, xend = xi, y = 0, yend = 1), alpha = 0.4, linetype = 2) +
  theme_light(base_size = 5) +
  scale_color_manual(values = rev(rcartocolor::carto_pal(5, "Safe"))[2:5]) +
  xlab("Natural logarithm of effect size") +
  ylab("Power")

@@

```{r polypower, fig.height = 3, fig.width = 12, fig.cap="..."}
patchwork::wrap_plots(append(plist, list(p), after = 0), 
                      design = map(1:11, ~patchwork::area(16, .x, 22, .x)) %>%
                        reduce(c) %>%
                        c(patchwork::area(1, 1, 15, 11), .))
```

@@

```{r}
p <- vi_survey %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "heteroskedasticity") %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "Visual test"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(data = mutate(heter_conv_sim, log_effect_size = log(effect_size)) %>%
                filter(x_dist == "uniform") %>%
                pivot_longer(F_p_value:SW_p_value) %>%
                mutate(name = gsub("_p_value", " test", name)), 
              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
              method = "glm", 
              method.args = list(family = binomial),
              se = FALSE) +
  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3)  +
  geom_segment(data = data.frame(xi = es), aes(x = xi, xend = xi, y = 0, yend = 1), alpha = 0.4, linetype = 2) +
  theme_light(base_size = 5) +
  scale_color_manual(values = rev(rcartocolor::carto_pal(5, "Safe"))[2:5]) +
  xlab("Natural logarithm of effect size") +
  ylab("Power")

@@

```{r heterpower, fig.height = 3, fig.width = 12, fig.cap="..."}
patchwork::wrap_plots(append(plist, list(p), after = 0), 
                      design = map(1:11, ~patchwork::area(16, .x, 22, .x)) %>%
                        reduce(c) %>%
                        c(patchwork::area(1, 1, 15, 11), .))
```

>>>>

Overall, the power analysis suggests that conventional tests do not match how humans perceive residual plots and thus determine model violations in these two scenarios. Non-severe model violations with human acceptable residual departures could often alarm by the conventional tests. This could limit the popularity of conventional tests in residual diagnostics among analysts.

@@

```{r}
p <- vi_survey %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "Visual test"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
                filter(x_dist == "uniform") %>%
                pivot_longer(F_p_value:SW_p_value) %>%
                mutate(name = gsub("_p_value", " test", name)), 
              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
              method = "glm", 
              method.args = list(family = binomial),
              se = FALSE) +
  geom_point(data = ~.x %>% 
               group_by(log_effect_size, reject_dirichlet) %>% 
               summarise(num_response = n()),
             aes(log_effect_size, reject_dirichlet, size = factor(num_response)), alpha = 0.3)  +
  geom_segment(data = data.frame(xi = es), aes(x = xi, xend = xi, y = 0, yend = 1), alpha = 0.4, linetype = 2) +
  theme_light(base_size = 5) +
  scale_color_manual(values = rev(rcartocolor::carto_pal(5, "Safe"))[2:5]) +
  scale_size_manual(values = c(1, 2, 3)) +
  xlab("Natural logarithm of effect size") +
  ylab("Power") +
  labs(col = "", size = "#lineups")

@@

```{r polypower, fig.height = 3, fig.width = 12, fig.cap="..."}
patchwork::wrap_plots(append(plist, list(p), after = 0), 
                      design = map(1:11, ~patchwork::area(16, .x, 22, .x)) %>%
                        reduce(c) %>%
                        c(patchwork::area(3, 1, 15, 11), .))
```

@@

```{r}
p <- vi_survey %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "heteroskedasticity") %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "Visual test"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(data = mutate(heter_conv_sim, log_effect_size = log(effect_size)) %>%
                filter(x_dist == "uniform") %>%
                pivot_longer(F_p_value:SW_p_value) %>%
                mutate(name = gsub("_p_value", " test", name)), 
              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
              method = "glm", 
              method.args = list(family = binomial),
              se = FALSE) +
    geom_point(data = ~.x %>% 
               group_by(log_effect_size, reject_dirichlet) %>% 
               summarise(num_response = n()),
             aes(log_effect_size, reject_dirichlet, size = factor(num_response)), alpha = 0.3)  +
  geom_segment(data = data.frame(xi = es), aes(x = xi, xend = xi, y = 0, yend = 1), alpha = 0.4, linetype = 2) +
  theme_light(base_size = 5) +
  scale_color_manual(values = rev(rcartocolor::carto_pal(5, "Safe"))[2:5]) +
  scale_size_manual(values = c(1, 2, 3)) +
  xlab("Natural logarithm of effect size") +
  ylab("Power") +
  labs(col = "", size = "#lineups")

@@

```{r heterpower, fig.height = 3, fig.width = 12, fig.cap="..."}
patchwork::wrap_plots(append(plist, list(p), after = 0), 
                      design = map(1:11, ~patchwork::area(16, .x, 22, .x)) %>%
                        reduce(c) %>%
                        c(patchwork::area(3, 1, 15, 11), .))
```

### 2022-11-29 paper_comparison/paper_comparison/paper_comparison.tex
finish power cmp

>>>>

Overall, the power analysis suggests that conventional tests do not
match how humans perceive residual plots and thus determine model
violations in these two scenarios. Non-severe model violations with
human acceptable residual departures could often alarm by the
conventional tests. This could limit the popularity of conventional
tests in residual diagnostics among analysts.
